---
title: "Exploratory_code"
author: "Sarah Roberts"
date: "2023-10-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I am trying to become familiar with the ways that we analyze acoustic data using a subset of Caroline's gabon data 

#1 Setup 

## 1.1 Load packages 
```{r}
devtools::install_github("https://github.com/DenaJGibbon/behaviouR") #this is Dena Clink's book of tutorials - https://bookdown.org/djc426/behaviouR-R-package-tutorials/ 
library(behaviouR)

library(ggfortify)
library(ggplot2)

```

## 1.2 read in data 
something is weird with the working directory so reading in manually
```{r}
WaveFile <- tuneR::readWave("~/Library/CloudStorage/Box-Box/innovations-research-tech/PAM/acoustic_BASS/data/subset/present1.wav")
WaveFile
```

double check the sampling rate 
```{r}
seewave::duration(WaveFile) * WaveFile@samp.rate
```

##1.3 plotting
Now we can plot the waveform from our sound file using the following code:
zoom in so that we can actually see the shape of the wave.

```{r}
seewave::oscillo(WaveFile, from = 0.1, to = 0.2)
seewave::oscillo(WaveFile, from = 0.18, to = 0.2)
```

Spectogram 
```{r}
SpectrogramSingle(sound.file = "~/Library/CloudStorage/Box-Box/innovations-research-tech/PAM/acoustic_BASS/data/subset/present1.wav")

#change the frequency 
SpectrogramSingle(sound.file = "~/Library/CloudStorage/Box-Box/innovations-research-tech/PAM/acoustic_BASS/data/subset/present1.wav", min.freq = 500, 
    max.freq = 2500,
    Colors = "Colors")
```

