---
title: "Exploratory_code"
author: "Sarah Roberts"
date: "2023-10-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I am trying to become familiar with the ways that we analyze acoustic data using a subset of Caroline's gabon data 

#1 Setup 

## 1.1 Load packages 
```{r}
devtools::install_github("https://github.com/DenaJGibbon/behaviouR") 

devtools::install_github("https://github.com/DenaJGibbon/gibbonR")
#this is Dena Clink's book of tutorials - https://bookdown.org/djc426/behaviouR-R-package-tutorials/ 
library(behaviouR)

library(ggfortify)
library(ggplot2)

install.packages("soundClass")

```

## 1.2 read in data 
something is weird with the working directory so reading in manually
```{r}
WaveFile <- tuneR::readWave("~/Library/CloudStorage/Box-Box/innovations-research-tech/PAM/acoustic_BASS/data/subset/present1.wav")
WaveFile
```

double check the sampling rate 
```{r}
seewave::duration(WaveFile) * WaveFile@samp.rate
```

##1.3 plotting
Now we can plot the waveform from our sound file using the following code:
zoom in so that we can actually see the shape of the wave.

```{r}
seewave::oscillo(WaveFile, from = 0.1, to = 0.2)
seewave::oscillo(WaveFile, from = 0.18, to = 0.2)
```

Spectogram 
```{r}
SpectrogramSingle(sound.file = "~/Library/CloudStorage/Box-Box/innovations-research-tech/PAM/acoustic_BASS/data/subset/present1.wav")

#change the frequency 
SpectrogramSingle(sound.file = "~/Library/CloudStorage/Box-Box/innovations-research-tech/PAM/acoustic_BASS/data/subset/present1.wav", min.freq = 500, 
    max.freq = 2500,
    Colors = "Colors")
```
#2 Model 

##2.1 model setup 
```{r}
MyFeatureDataFrame <- MFCCFunction(input.dir = "/Users/sarahroberts/Library/CloudStorage/Box-Box/innovations-research-tech/PAM/acoustic_BASS/data/subset",min.freq = 500,max.freq=2000)
```

```{r}
pca_res <- prcomp(MyFeatureDataFrame[,-c(1)], scale. = TRUE)

MyFeatureDataFrame$class2 <- c("absent", "absent", "absent", "absent", "absent", "present", "present", "present", "present", "present")
ggplot2::autoplot(pca_res, data = MyFeatureDataFrame,
         colour = 'class2')

```


```{r}

trainingdata <- MyFeatureDataFrame
trainingdata$class <- as.factor(trainingdata$class2)

trainingdata <- trainingdata %>% dplyr::select(-class2)

ml.model.rf <- randomForest::randomForest(x=trainingdata[, 2:ncol(trainingdata)], y = trainingdata$class)


print(ml.model.rf)

ml.model.svm <- e1071::svm(trainingdata[, 2:ncol(trainingdata)], trainingdata$class, kernel = "radial", 
                           cross = 3,
                           probability = TRUE)

print(paste('SVM accuracy',ml.model.svm$tot.accuracy))

print(ml.model.svm)
```

